---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Repository overview

This repository contains the code accompanying the paper
**"Collective Outlier Detection and Enumeration with Conformalized Closed Testing."**

It provides implementations of ACODE (Automatic Conformal Outlier Detection and Enumeration), closed-testing procedures based on rank tests, and all simulation and real-data experiments reported in the paper, including those in the appendices. The repository includes:

- An R package (`nout/`) implementing ACODE and the closed-testing procedures.
- Python and R experiment scripts (`experiments/`) used to run the simulation studies and real-data analyses, together with shell launchers that drive parameter sweeps on either a Slurm cluster or a single machine.
- Core Python utilities (`methods/`) for data generation, conformal scoring, ACODE orchestration, and interfacing with the R package for closed testing via `rpy2`.
- R scripts (`experiments/figures_and_tables/`) that assemble the final figures and tables in the paper from the experiment outputs.
- A metadata layer (`experiments/metadata/`) mapping every figure and table in the manuscript to the scripts that produce it (see the *Reproducibility map* section below).
- Precomputed critical value tables (`tables/`) for commonly used local tests (Fisher, Wilcoxon–Mann–Whitney, and Shiraishi tests against Lehmann alternatives), used to accelerate small-sample calculations where asymptotic approximations may be unreliable.
- A `data/` folder containing the datasets used in the experiments (see the *Data* section below).

The sections below describe how the repository is organized, how to install the dependencies, and how to reproduce every figure and table reported in the paper (including those in the appendices).

# Data

> **Note.** The `data/` folder is not hosted on GitHub due to size and possible licensing restrictions. All processed data files are included in the supplementary material. Original sources are listed below; see Appendix A6.3 of the manuscript for further details.

| Dataset | File(s) | Description | Source / Prior usage |
|----|----|----|----|
| **LHC Olympics 2020 (LHCO)** | `events_anomalydetection_v2.features.h5`, `events_anomalydetection_Z_XY_qqq.features.h5`, `events_anomalydetection_Z_XY_qqq.h5` | Community benchmark for anomaly detection in high-energy physics. | Kasieczka et al., 2021. |
| **ALOI** | `aloi.arff` | 27 variables, 1,508 outliers, 48,026 inliers. | Amsterdam Library of Object Images (Geusebroek et al., 2005). Previously used in Bates et al., 2023. |
| **Covertype** | `cover.mat` | 10 variables, 2,747 outliers, 286,048 inliers. | Covertype dataset from UCI / ODDS repository. Previously used in Bates et al., 2023. |
| **CreditCard** | `creditcard.csv` | 30 variables, 492 outliers, 284,315 inliers. | Kaggle Credit Card Fraud Dataset. Used in Bates et al., 2023; and Marandon et al., 2024. |
| **Mammography** | `mammography.mat` | 6 variables, 260 outliers, 10,923 inliers. | Mammography dataset from ODDS repository. Used in Bates et al., 2023; Liang et al., 2024; Marandon et al., 2024. |
| **Pendigits** | `pendigits.mat` | 16 variables, 156 outliers, 6,714 inliers. | Pen-Based Recognition of Handwritten Digits dataset (ODDS repository). Used in Bates et al., 2023. |
| **Shuttle** | `shuttle.mat` | 9 variables, 3,511 outliers, 45,586 inliers. | Statlog Shuttle dataset (ODDS repository). Used in Bates et al., 2023; and Marandon et al., 2024. |

# Code organization

## R package `nout/` (required)

All closed-testing procedures rely on the R package **`nout`**, included in this repository under `nout/`. This package is called from both the R experiments and the Python experiments (via `rpy2`), so installing `nout` is required for all workflows.

Before running experiments, install the package locally:

~~~r
install.packages("devtools")  # if needed
devtools::install("nout/")
~~~

## Methods (`methods/`)

Core utilities used by the Python experiments: data handling, synthetic data generators, conformal score construction, the ACODE implementation, and the interface to the R package for closed testing.

- `methods/models.py` — synthetic data generators used in the simulation studies.
- `methods/my_utils.py` — utilities for experiment bookkeeping and evaluation.
- `methods/conformal.py` — ACODE implementation. Imports the R package `nout` via `rpy2` (`importr("nout")`) and exposes the closed-testing routines for use inside the Python code.

## Experiments (`experiments/`)

This is where the simulation and real-data experiments are launched and where their outputs land. The folder layout is:

~~~
experiments/
├── exp_A.py / exp_A.sh                          # Python experiment driver (synthetic, real tabular, LHCO)
├── exp_B1_global_testing.R / .sh                # R experiment: global testing (Appendix B1)
├── exp_B2_enumeration.R    / .sh                # R experiment: enumeration (Appendix B2)
├── exp_C1_time.R                                # R experiment: timing comparison (Appendix C1)
├── exp_C2_simes.R                               # R experiment: Simes-permutation comparison (Appendix C2)
├── make_figures_tables.sh                       # builds all figures and tables sequentially
├── submit_exp_A.sh                              # Grid launcher for exp_A (for Figures 1-4, Figures A2-A10, Figures A13-A22, Table A4)
├── submit_exp_B1.sh                             # Grid launcher for exp_B1 (for Figure A11)
├── submit_exp_B2.sh                             # Grid launcher for exp_B2 (for Figure A12)
├── submit_exp_C1.sh                             # Wrapper for exp_C1 (for Figure A1)
├── submit_exp_C2.sh                             # Wrapper for exp_C2 (for Table A2)
├── run_all_experiments.sh                       # Meta-launcher: submits every experiment via map_submit.tsv
├── figures_and_tables/                          # R scripts that turn experiment outputs into final figures/tables
│   ├── make_fig*.R, make_tab*.R                 # one make_*.R per paper figure/table group
│   ├── utils_plotting_*.R                       # helper functions for making figures and tables
│   ├── figures/                                 # produced figures (.pdf)
│   ├── tables/                                  # produced tables (.tex)
│   └── logs/                                    # per-script R output logs
├── metadata/                                    # provenance map: paper item ↔ scripts ↔ result directory
│   ├── map_full.tsv, map_full.md, map_full.tex  # consolidated map (see Reproducibility map section)
│   ├── map_submit.tsv, map_results.tsv          # extracted from submit_*.sh headers
│   ├── map_scripts.tsv, map_figures.tsv         # extracted from make_*.R headers
│   ├── extract_submit_metadata.sh               # rebuilds the submit-side maps
│   ├── extract_make_metadata.sh                 # rebuilds the make-side maps
│   └── cross_reference_metadata.sh              # joins everything into map_full.tsv / .md
├── logs/                                        # cluster job logs (one subdir per result key)
└── results/                                     # experiment outputs (one subdir per result key)
~~~


All submit launchers accept the same flags for consistency:

- `--cluster` (default) submits jobs via `sbatch`.
- `--local` runs jobs sequentially on the current machine.
- `--dry-run` prints the commands without executing or submitting them.
- `--force` re-runs experiments even if result files already exist.

They also skip any configuration whose output already exists, so launching the same script a second time only fills in the missing pieces.

# Reproducibility map (paper item → scripts)

```{r reproducibility-map, echo=FALSE, results='asis'}
cat(readLines("experiments/metadata/map_full.md"), sep = "\n")
```

~~~bash
cd experiments/metadata
./extract_submit_metadata.sh
./extract_make_metadata.sh
./cross_reference_metadata.sh
~~~

This is also the table to consult when only a subset of figures needs to be regenerated.

# Reproducibility workflow

## Dependencies

```{r dependencies, echo=FALSE, results='asis'}
cat(readLines("metadata/dependencies.md"), sep = "\n")
```

## Workflow

```{r workflow, echo=FALSE, results='asis'}
cat(readLines("metadata/workflow.md"), sep = "\n")
```


## Computing time

Each individual experiment runs in seconds to a few minutes on a standard laptop. The aggregate computational cost arises from the large number of configurations and repetitions, not from any single run. On a Slurm cluster with dozens of parallel cores, the full pipeline completes in a few hours; on a single laptop, completing it sequentially is impractical.