---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nout

<!-- badges: start -->
<!-- badges: end -->

The R package ***nout*** implements the closed testing shortcuts used by ACODE. 

## Installation

``` r
# install.packages("devtools")
devtools::install_github("msesia/conformal_nout", subdir="nout")
```

## Example

This is a basic example on synthetic data, where inlier scores are generated from a Uniform(0,1) and outlier scores are generated under Lehmann's alternative of order k=2. All observations in the test set are outliers.


```{r example}
library(nout)

set.seed(321)

# Auxiliary functions to generate outliers scores
g2 = function(x, k=2) ifelse(x<1 & x>0, k*x^(k-1), 0)
rg2 = function(rnull, k=2) max(rnull(k))

# Generating calibration and test score vectors
# Outlier distribution is the Lehmann's alternative of order k=2
X = runif(20)
Y = replicate(20, rg2(rnull=runif))

```


### Lower bound without selection

```{r}
# Confidence lower-bound for the number of outliers using Closed Testing with local Simes test
res1 = d_selection_simes(X, Y,alpha = 0.1)
res1

# Confidence lower-bound for the number of outliers using Closed Testing with local Fisher's method
res2 = d_selection_fisher(X, Y, B=100, n_perm=0, alpha=0.1)
res2

# Confidence lower-bound for the number of outliers using Closed Testing with local Wilcoxon sum-rank test 
res3 = d_selection_higher(X, Y, local_test="WMW", n_perm=0, B=100, alpha=0.1)
res3

# Confidence lower-bound for the number of outliers using Closed Testing with local higher order
# Wilcoxon sum-rank test of order k=2
res4 = d_selection_higher(X, Y, local_test="higher", k=2, n_perm=0, B=100, alpha=0.1)
res4

# Confidence lower-bound for the number of outliers using Closed Testing with local oracle Shiraishi test
res5 = d_selection_G(X, Y, g.hat = g2, monotone=TRUE, B=100, alpha=0.1)
res5

```


### Lower bound with selection

The selection set is randomly chosen among the test index set.

```{r }
S = sample(1:20, size = 16)

# Confidence lower-bound for the number of outliers using Closed Testing with local Simes test
resS1 = d_selection_simes(X, Y, S=S, alpha = 0.1)
resS1

# Confidence lower-bound for the number of outliers using Closed Testing with local Fisher's method
resS2 = d_selection_fisher(X, Y, S=S, B=100, n_perm=0, alpha=0.1)
resS2

# Confidence lower-bound for the number of outliers using Closed Testing with local Wilcoxon sum-rank test 
resS3 = d_selection_higher(X, Y, S=S, local_test="WMW", n_perm=0, B=100, alpha=0.1)
resS3

# Confidence lower-bound for the number of outliers using Closed Testing with local higher order
# Wilcoxon sum-rank test of order k=2
resS4 = d_selection_higher(X, Y, S=S, local_test="higher", k=2, n_perm=0, B=100, alpha=0.1)
resS4

# Confidence lower-bound for the number of outliers using Closed Testing with local oracle Shiraishi test
resS5 = d_selection_G(X, Y, S=S, g.hat = g2, monotone=TRUE, B=100, alpha=0.1)
resS5

```
